---
title: "ECO395M STAT LEARNING Homework 1" 
author: "Mingwei Li, Xinyu Leng, Hongjin Long"
thanks: "Mingwei Li, Xinyu Leng and Hongjin Long are master students of economics, The University of Texas at Austin"
output:
  pdf_document: 
    number_sections: yes
  html_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\begin{abstract}
This document is the first homework of ECO395M STAT LEARNING. The projects \textbf{Gas Price}, \textbf{a bike share network}, \textbf{flights at ABIA} and \textbf{K-nearest neighbors} are included in this document. The whole project is available at \textcolor{blue}{\href{https://github.com/mliw/datamining_hk1}{here}}. 
\begin{figure}[!htbp]
	\centering
	\includegraphics[width=0.85\textwidth]{pic/0.jpg}
\end{figure}

\end{abstract}

\newpage
\tableofcontents

\newpage
\section{Gas Price}
```{r, echo=FALSE}
gasprice = read.csv("data/GasPrices.csv")
library(ggplot2)
```

\textbf{(A) Gas stations charge more if they lack direct competition in sight (boxplot).}

```{r fig1, fig.width=4, fig.align = "center",echo=TRUE}
p0 = ggplot(data=gasprice) + 
  geom_boxplot(aes(x=Competitors, y=Price))
p0
```
\textbf{Claim}: Gas stations charge more if they lack direct competition in sight

\textbf{Conclusion}: The theory is supported by the data. As the left box is higher than the right box.

\newpage
\textbf{(B) The richer the area, the higher the gas price (scatter plot).}

```{r fig2, fig.width=4, fig.align = "center",echo=TRUE}
p0 = ggplot(data=gasprice) + 
  geom_point(aes(x=Income, y=Price))
p0
```
\textbf{Claim}: Gas stations charge more if they lack direct competition in sight

\textbf{Conclusion}: The theory is supported by the data. As the left box is higher than the right box.

